Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-06-29 paper/AlignedIsotropicGW.tex
Moved bibliographies into text, split between main and methods sections.

@@

\begin{addendum}
\item We thank Richard O'Shaughnessy, Christopher Berry, Davide
  Gerosa, and Salvatore Vitale for discussions and comments on this
  work.  WF, SS, IM and AV were supported in part by the STFC.  MCM
  acknowledges support of the University of Birmingham Institute for
  Advanced Study Distinguished Visiting Fellows program.  SS and IM
  acknowledge support from the National Science Foundation under Grant
  No. NSF PHY11-25915.
\item[Supplementary Information] Supplementary Information is linked
  to the online version of the paper at
  \url{http://www.nature.com/nature}.
\item[Author Contributions] All authors contributed to the work
  presented in this paper.
\item[Author Information] Reprints and permissions information is
  available at \url{http://www.nature.com/reprints}.  The authors
  declare no competing financial interests.  Correspondence and
  requests for materials should be addressed to
  \href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
\end{addendum}

@@

\bibliographystyle{naturemag-arxiv}
\bibliography{AlignedIsotropicGW}

>>>>

\newcounter{mainbib}

@@

\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem{2016PhRvL.116f1102A}
\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{{Observation of Gravitational Waves from a Binary
  Black Hole Merger}}.
\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{061102}
  (\bibinfo{year}{2016}).

\bibitem{2016PhRvL.116x1103A}
\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{{GW151226: Observation of Gravitational Waves from a
  22-Solar-Mass Binary Black Hole Coalescence}}.
\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{241103}
  (\bibinfo{year}{2016}).

\bibitem{O1-BBH}
\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{{Binary Black Hole Mergers in the First Advanced LIGO
  Observing Run}}.
\newblock \emph{\bibinfo{journal}{Physical Review X}}
  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{041015} (\bibinfo{year}{2016}).

\bibitem{PhysRevLett.118.221101}
\bibinfo{author}{Abbott, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{Gw170104: Observation of a 50-solar-mass binary black
  hole coalescence at redshift 0.2}.
\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
  \textbf{\bibinfo{volume}{118}}, \bibinfo{pages}{221101}
  (\bibinfo{year}{2017}).
\newblock
  \urlprefix\url{https://link.aps.org/doi/10.1103/PhysRevLett.118.221101}.

\bibitem{SigurdssonHernquist:1993}
\bibinfo{author}{{Sigurdsson}, S.} \& \bibinfo{author}{{Hernquist}, L.}
\newblock \bibinfo{title}{{Primordial black holes in globular clusters}}.
\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{364}},
  \bibinfo{pages}{423--425} (\bibinfo{year}{1993}).

\bibitem{PZMcMillan:2000}
\bibinfo{author}{{Portegies Zwart}, S.~F.} \& \bibinfo{author}{McMillan, S.
  L.~W.}
\newblock \bibinfo{title}{Black hole mergers in the universe}.
\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{528}},
  \bibinfo{pages}{L17} (\bibinfo{year}{2000}).

\bibitem{Rodriguez:2015}
\bibinfo{author}{{Rodriguez}, C.~L.} \emph{et~al.}
\newblock \bibinfo{title}{{Binary Black Hole Mergers from Globular Clusters:
  Implications for Advanced LIGO}}.
\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
  \textbf{\bibinfo{volume}{115}}, \bibinfo{pages}{051101}
  (\bibinfo{year}{2015}).

\bibitem{Stone:2016}
\bibinfo{author}{{Stone}, N.~C.}, \bibinfo{author}{{Metzger}, B.~D.} \&
  \bibinfo{author}{{Haiman}, Z.}
\newblock \bibinfo{title}{{Assisted inspirals of stellar mass black holes
  embedded in AGN discs: solving the `final au problem'}}.
\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{464}},
  \bibinfo{pages}{946--954} (\bibinfo{year}{2017}).

\bibitem{2016ApJ...832L...2R}
\bibinfo{author}{{Rodriguez}, C.~L.}, \bibinfo{author}{{Zevin}, M.},
  \bibinfo{author}{{Pankow}, C.}, \bibinfo{author}{{Kalogera}, V.} \&
  \bibinfo{author}{{Rasio}, F.~A.}
\newblock \bibinfo{title}{{Illuminating Black Hole Binary Formation Channels
  with Spins in Advanced LIGO}}.
\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{832}},
  \bibinfo{pages}{L2} (\bibinfo{year}{2016}).

\bibitem{TutukovYungelson:1993}
\bibinfo{author}{Tutukov, A.~V.} \& \bibinfo{author}{Yungelson, L.~R.}
\newblock \bibinfo{title}{The merger rate of neutron star and black hole
  binaries}.
\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
  Society}} \textbf{\bibinfo{volume}{260}}, \bibinfo{pages}{675--678}
  (\bibinfo{year}{1993}).

\bibitem{2016Natur.534..512B}
\bibinfo{author}{{Belczynski}, K.}, \bibinfo{author}{{Holz}, D.~E.},
  \bibinfo{author}{{Bulik}, T.} \& \bibinfo{author}{{O'Shaughnessy}, R.}
\newblock \bibinfo{title}{{The first gravitational-wave source from the
  isolated evolution of two stars in the 40-100 solar mass range}}.
\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{534}},
  \bibinfo{pages}{512--515} (\bibinfo{year}{2016}).

\bibitem{Stevenson:2017}
\bibinfo{author}{{Stevenson}, S.} \emph{et~al.}
\newblock \bibinfo{title}{{Formation of the first three gravitational-wave
  observations through isolated binary evolution}}.
\newblock \emph{\bibinfo{journal}{Nat. Commun.}} \textbf{\bibinfo{volume}{8}},
  \bibinfo{pages}{14906} (\bibinfo{year}{2017}).

\bibitem{MandeldeMink:2016}
\bibinfo{author}{{Mandel}, I.} \& \bibinfo{author}{{de Mink}, S.~E.}
\newblock \bibinfo{title}{{Merging binary black holes formed through chemically
  homogeneous evolution in short-period stellar binaries}}.
\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{458}},
  \bibinfo{pages}{2634--2647} (\bibinfo{year}{2016}).

\bibitem{Marchant:2016}
\bibinfo{author}{{Marchant}, P.}, \bibinfo{author}{{Langer}, N.},
  \bibinfo{author}{{Podsiadlowski}, P.}, \bibinfo{author}{{Tauris}, T.~M.} \&
  \bibinfo{author}{{Moriya}, T.~J.}
\newblock \bibinfo{title}{{A new route towards merging massive black holes}}.
\newblock \emph{\bibinfo{journal}{\aap}} \textbf{\bibinfo{volume}{588}},
  \bibinfo{pages}{A50} (\bibinfo{year}{2016}).

\bibitem{2016ApJ...833L...1A}
\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{{The Rate of Binary Black Hole Mergers Inferred from
  Advanced LIGO Observations Surrounding GW150914}}.
\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{833}},
  \bibinfo{pages}{L1} (\bibinfo{year}{2016}).

\bibitem{2015PhR...548....1M}
\bibinfo{author}{{Miller}, M.~C.} \& \bibinfo{author}{{Miller}, J.~M.}
\newblock \bibinfo{title}{{The masses and spins of neutron stars and
  stellar-mass black holes}}.
\newblock \emph{\bibinfo{journal}{\physrep}} \textbf{\bibinfo{volume}{548}},
  \bibinfo{pages}{1--34} (\bibinfo{year}{2015}).

\bibitem{2017CQGra..34cLT01V}
\bibinfo{author}{{Vitale}, S.}, \bibinfo{author}{{Lynch}, R.},
  \bibinfo{author}{{Sturani}, R.} \& \bibinfo{author}{{Graff}, P.}
\newblock \bibinfo{title}{{Use of gravitational waves to probe the formation
  channels of compact binaries}}.
\newblock \emph{\bibinfo{journal}{Classical and Quantum Gravity}}
  \textbf{\bibinfo{volume}{34}}, \bibinfo{pages}{03LT01}
  (\bibinfo{year}{2017}).

\bibitem{1999MNRAS.305..654K}
\bibinfo{author}{{King}, A.~R.} \& \bibinfo{author}{{Kolb}, U.}
\newblock \bibinfo{title}{{The evolution of black hole mass and angular
  momentum}}.
\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{305}},
  \bibinfo{pages}{654--660} (\bibinfo{year}{1999}).

\bibitem{2015ApJ...800...17F}
\bibinfo{author}{{Fragos}, T.} \& \bibinfo{author}{{McClintock}, J.~E.}
\newblock \bibinfo{title}{{The Origin of Black Hole Spin in Galactic Low-mass
  X-Ray Binaries}}.
\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{800}},
  \bibinfo{pages}{17} (\bibinfo{year}{2015}).

\bibitem{2016MNRAS.462..844K}
\bibinfo{author}{{Kushnir}, D.}, \bibinfo{author}{{Zaldarriaga}, M.},
  \bibinfo{author}{{Kollmeier}, J.~A.} \& \bibinfo{author}{{Waldman}, R.}
\newblock \bibinfo{title}{{GW150914: spin-based constraints on the merger time
  of the progenitor system}}.
\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{462}},
  \bibinfo{pages}{844--849} (\bibinfo{year}{2016}).

\bibitem{2017arXiv170200885Z}
\bibinfo{author}{{Zaldarriaga}, M.}, \bibinfo{author}{{Kushnir}, D.} \&
  \bibinfo{author}{{Kollmeier}, J.~A.}
\newblock \bibinfo{title}{{Research Note: The Expected Spins of Gravitational
  Wave Sources With Isolated Field Binary Progenitors}}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
\newblock \eprint{arXiv:1702.00885}.

\bibitem{2017arXiv170203952H}
\bibinfo{author}{{Hotokezaka}, K.} \& \bibinfo{author}{{Piran}, T.}
\newblock \bibinfo{title}{{Implications of the low binary black hole spins
  observed by LIGO}}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
\newblock \eprint{arXiv:1702.03952}.

\bibitem{Albrecht:2009}
\bibinfo{author}{{Albrecht}, S.}, \bibinfo{author}{{Reffert}, S.},
  \bibinfo{author}{{Snellen}, I.~A.~G.} \& \bibinfo{author}{{Winn}, J.~N.}
\newblock \bibinfo{title}{{Misaligned spin and orbital axes cause the anomalous
  precession of DI Herculis}}.
\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{461}},
  \bibinfo{pages}{373--376} (\bibinfo{year}{2009}).

\bibitem{Martin:2008b}
\bibinfo{author}{{Martin}, R.~G.}, \bibinfo{author}{{Reis}, R.~C.} \&
  \bibinfo{author}{{Pringle}, J.~E.}
\newblock \bibinfo{title}{{Misalignment of the microquasar V4641 Sgr (SAX
  J1819.3-2525)}}.
\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{391}},
  \bibinfo{pages}{L15--L18} (\bibinfo{year}{2008}).

\bibitem{MorningstarMiller:2014}
\bibinfo{author}{{Morningstar}, W.~R.} \& \bibinfo{author}{{Miller}, J.~M.}
\newblock \bibinfo{title}{{The Spin of the Black Hole 4U 1543-47}}.
\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{793}},
  \bibinfo{pages}{L33} (\bibinfo{year}{2014}).

\bibitem{2000ApJ...541..319K}
\bibinfo{author}{{Kalogera}, V.}
\newblock \bibinfo{title}{{Spin-Orbit Misalignment in Close Binaries with Two
  Compact Objects}}.
\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{541}},
  \bibinfo{pages}{319--328} (\bibinfo{year}{2000}).

\bibitem{2011ApJ...742...81F}
\bibinfo{author}{{Farr}, W.~M.}, \bibinfo{author}{{Kremer}, K.},
  \bibinfo{author}{{Lyutikov}, M.} \& \bibinfo{author}{{Kalogera}, V.}
\newblock \bibinfo{title}{{Spin Tilts in the Double Pulsar Reveal Supernova
  Spin Angular-momentum Production}}.
\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{742}},
  \bibinfo{pages}{81} (\bibinfo{year}{2011}).

\bibitem{2015PhRvD..92f4016G}
\bibinfo{author}{{Gerosa}, D.}, \bibinfo{author}{{Kesden}, M.},
  \bibinfo{author}{{Sperhake}, U.}, \bibinfo{author}{{Berti}, E.} \&
  \bibinfo{author}{{O'Shaughnessy}, R.}
\newblock \bibinfo{title}{{Multi-timescale analysis of phase transitions in
  precessing black-hole binaries}}.
\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{92}},
  \bibinfo{pages}{064016} (\bibinfo{year}{2015}).

\bibitem{2004PhRvD..70l4020S}
\bibinfo{author}{{Schnittman}, J.~D.}
\newblock \bibinfo{title}{{Spin-orbit resonance and the evolution of compact
  binary systems}}.
\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{70}},
  \bibinfo{pages}{124020} (\bibinfo{year}{2004}).

\bibitem{2007ApJ...661L.147B}
\bibinfo{author}{{Bogdanovi{\'c}}, T.}, \bibinfo{author}{{Reynolds}, C.~S.} \&
  \bibinfo{author}{{Miller}, M.~C.}
\newblock \bibinfo{title}{{Alignment of the Spins of Supermassive Black Holes
  Prior to Coalescence}}.
\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{661}},
  \bibinfo{pages}{L147--L150} (\bibinfo{year}{2007}).

\setcounter{mainbib}{\value{enumiv}}
\end{thebibliography}

\begin{addendum}
\item We thank Richard O'Shaughnessy, Christopher Berry, Davide
  Gerosa, and Salvatore Vitale for discussions and comments on this
  work.  WF, SS, IM and AV were supported in part by the STFC.  MCM
  acknowledges support of the University of Birmingham Institute for
  Advanced Study Distinguished Visiting Fellows program.  SS and IM
  acknowledge support from the National Science Foundation under Grant
  No. NSF PHY11-25915.
\item[Author Contributions] All authors contributed to the work
  presented in this paper.
\item[Author Information] Reprints and permissions information is
  available at \url{http://www.nature.com/reprints}.  The authors
  declare no competing financial interests.  Correspondence and
  requests for materials should be addressed to
  \href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
\end{addendum}

@@

\begin{thebibliography}{10}
\setcounter{enumiv}{\value{mainbib}}

\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem{Julia}
\bibinfo{author}{Bezanson, J.}, \bibinfo{author}{Karpinski, S.},
  \bibinfo{author}{Shah, V.~B.} \& \bibinfo{author}{Edelman, A.}
\newblock \bibinfo{title}{{J}ulia: A fast dynamic language for technical
  computing}  (\bibinfo{year}{2012}).
\newblock \eprint{arXiv:1209.5145}.

\bibitem{NumPy}
\bibinfo{author}{van~der Walt, S.}, \bibinfo{author}{Colbert, S.~C.} \&
  \bibinfo{author}{Varoquaux, G.}
\newblock \bibinfo{title}{The numpy array: A structure for efficient numerical
  computation}.
\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{22--30}
  (\bibinfo{year}{2011}).

\bibitem{SciPy}
\bibinfo{author}{Jones, E.}, \bibinfo{author}{Oliphant, T.},
  \bibinfo{author}{Peterson, P.} \emph{et~al.}
\newblock \bibinfo{title}{{SciPy}: Open source scientific tools for {Python}}
  (\bibinfo{year}{2001--}).
\newblock \urlprefix\url{http://www.scipy.org/}.
\newblock \bibinfo{note}{[Online; accessed 24 May 2017]}.

\bibitem{Matplotlib}
\bibinfo{author}{Hunter, J.~D.}
\newblock \bibinfo{title}{Matplotlib: A 2d graphics environment}.
\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{90--95} (\bibinfo{year}{2007}).

\bibitem{IPython}
\bibinfo{author}{Pérez, F.} \& \bibinfo{author}{Granger, B.~E.}
\newblock \bibinfo{title}{Ipython: A system for interactive scientific
  computing}.
\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{21--29} (\bibinfo{year}{2007}).
\newblock
  \urlprefix\url{http://aip.scitation.org/doi/abs/10.1109/MCSE.2007.53}.

\bibitem{2016PhRvL.116x1102A}
\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{{Properties of the Binary Black Hole Merger
  GW150914}}.
\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{241102}
  (\bibinfo{year}{2016}).

\bibitem{Stevenson:2017spin}
\bibinfo{author}{{Stevenson}, S.}, \bibinfo{author}{{Berry}, C.~P.~L.} \&
  \bibinfo{author}{{Mandel}, I.}
\newblock \bibinfo{title}{{Hierarchical analysis of gravitational-wave
  measurements of binary black hole spin-orbit misalignments}}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
\newblock \eprint{1703.06873}.

\bibitem{2015PhRvD..91d2003V}
\bibinfo{author}{{Veitch}, J.} \emph{et~al.}
\newblock \bibinfo{title}{{Parameter estimation for compact binaries with
  ground-based gravitational-wave observations using the LALInference software
  library}}.
\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{91}},
  \bibinfo{pages}{042003} (\bibinfo{year}{2015}).

\bibitem{2010ApJ...725.2166H}
\bibinfo{author}{{Hogg}, D.~W.}, \bibinfo{author}{{Myers}, A.~D.} \&
  \bibinfo{author}{{Bovy}, J.}
\newblock \bibinfo{title}{{Inferring the Eccentricity Distribution}}.
\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{725}},
  \bibinfo{pages}{2166--2175} (\bibinfo{year}{2010}).

\bibitem{2010PhRvD..81h4029M}
\bibinfo{author}{{Mandel}, I.}
\newblock \bibinfo{title}{{Parameter estimation on gravitational waves from
  multiple coalescing binaries}}.
\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{81}},
  \bibinfo{pages}{084029} (\bibinfo{year}{2010}).

\bibitem{2017arXiv170306869F}
\bibinfo{author}{{Fishbach}, M.}, \bibinfo{author}{{Holz}, D.} \&
  \bibinfo{author}{{Farr}, B.}
\newblock \bibinfo{title}{{Are LIGO's Black Holes Made From Smaller Black
  Holes?}}
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
\newblock \eprint{arXiv:1703.06869}.

\bibitem{2017arXiv170306223G}
\bibinfo{author}{{Gerosa}, D.} \& \bibinfo{author}{{Berti}, E.}
\newblock \bibinfo{title}{{First or second generation? Black hole census with
  gravitational wave observations}}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
\newblock \eprint{arXiv:1703.06223}.

\bibitem{2016PhRvD..93l2003A}
\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{{GW150914: First results from the search for binary
  black hole coalescence with Advanced LIGO}}.
\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{93}},
  \bibinfo{pages}{122003} (\bibinfo{year}{2016}).

\bibitem{2016arXiv161107531T}
\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{{Effects of waveform model systematics on the
  interpretation of GW150914}}.
\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2016}).
\newblock \eprint{arXiv:1611.07531}.

\bibitem{2016PhRvD..94f4035A}
\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
\newblock \bibinfo{title}{{Directly comparing GW150914 with numerical solutions
  of Einstein's equations for binary black hole coalescence}}.
\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{94}},
  \bibinfo{pages}{064035} (\bibinfo{year}{2016}).

\end{thebibliography}

%\bibliographystyle{naturemag-arxiv}
%\bibliography{AlignedIsotropicGW}
